EEG2fMRI: Cross-Modal Synthesis for Functional Neuroimaging

### 2020-11-19 article/article.tex
Update introduction

\title{EEG2fMRI: Cross-Modal Synthesis for Functional Neuroimaging}
\author{Paul Bricman}
\author{Jelmer Borst}
\affil{University of Groningen}
\date{July 2020}

@@

Neural activity is central to investigating cognition. It has been employed to study the neural correlates of numerous cognitive functions, such as perception, attention, and memory. It has also been employed for characterizing, diagnosing, and ameliorating a wide range of neurological conditions.

@@

Given the lack of techniques which individually exhibit high overall performance, there have been numerous attempts to integrate their complementary qualities by using multiple techniques simultaneously. EEG-informed fMRI, fMRI-informed EEG, and neurogenerative modelling are all methods of integrating data from simultaneous EEG and fMRI recordings. However, these methods require two techniques and are limited to niched applications.

@@

In addition to uni-modal synthesis, investigators have also proposed cross-modal synthesis for augmenting neuroimaging techniques. This task consists of reconstructing data collected through a target modality based on data collected through a distinct source modality, by exploiting the inherent relation between them. In a setting where both source and target modalities are available, this task might have limited utility. However, the value of cross-modal synthesis comes from settings where only one modality is available. In such settings, data collected through an unavailable modality can be approximated. Unfortunately, this approach has largely been used for augmenting structural neuroimaging, with limited analogous efforts for functional neuroimaging.

Despite the major demand for high-performance functional neuroimaging and the success of cross-modal synthesis for structural neuroimaging, limited attention has been given to cross-modal synthesis for functional neuroimaging. The purpose of the current work is to formalize and explore the feasibility of EEG-fMRI synthesis. This comes as a natural extension of the growing collection of functional neuroimaging augmentation methods.

The task of modelling the relation between two modalities can be addressed through different approaches.

>>>>

\title{EEG2fMRI: Cross-Modal Synthesis for Functional Neuroimaging}
\author{Paul Bricman}
\author{Jelmer Borst}
\affil{University of Groningen}
\date{July 2021}

@@

Neural activity is central to investigating cognition. In cognitive neuroscience, it has been employed to study the neural correlates of numerous cognitive functions, such as perception, attention, and memory. In cognitive modelling, neural activity has been used to inform the development of a wide range of cognitive architectures. In clinical neuroscience, it has been employed for characterizing, diagnosing, and ameliorating a wide range of neurological conditions.

@@

Given the lack of techniques which individually exhibit high overall performance, there have been numerous attempts to integrate their complementary qualities by using multiple techniques simultaneously. EEG-informed fMRI, fMRI-informed EEG, and neurogenerative modelling are all methods of integrating data from simultaneous EEG and fMRI recordings. However, these methods require access to two techniques and are limited to niched applications.

@@

In addition to uni-modal synthesis, investigators have also proposed cross-modal synthesis for augmenting neuroimaging techniques. This task consists of reconstructing data collected through a target modality based on data collected through a distinct source modality, by exploiting the inherent relation between them. In a setting where both source and target modalities are available, this task might have limited utility. However, the value of cross-modal synthesis comes from settings where only one modality is available. In such settings, existing data can be used to approximate data collected through an unavailable modality. Unfortunately, this approach has largely been used for augmenting structural neuroimaging, with limited analogous efforts for functional neuroimaging.

Despite the major demand for high-performance functional neuroimaging and the success of cross-modal synthesis for structural neuroimaging, limited attention has been given to cross-modal synthesis for functional neuroimaging. The purpose of the current work is to formalize and explore the feasibility of this method. This comes as a natural extension to the growing collection of functional neuroimaging augmentation methods.

The task of modelling the relation between two modalities can be addressed through at least three different approaches. First, if simultaneous multi-modal recordings are available, supervised learning can be used to model the mapping between the two domains based on paired samples. Second, if simultaneous multi-modal recordings are not available, but separate recordings of analogous mental states with different modalities are available (i.e. two recording sessions based on the same task but using different techniques), semi-supervised learning can be used to model the mapping through a task-related signal. This is analogous to the task of machine translation, which is based on a cross-lingual signal. Third, if few such recordings exist, unsupervised learning can be used to learn the mapping by aligning latent spaces associated with the two modalities. This is analogous to machine translation where no paired multilingual corpora exist.
